Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Utilizing Expert Knowledge in Model Modification

### 2024-08-06 paper/main.tex
Updates paper

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Expert-In-The-Loop Causal Discovery}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
	Causal Discovery has received substantial attention in the DAG
	literature and many automated algorithms have been proposed for
	constructing DAGs from data. However, the application of these
	algorithms in applied fields remain limited due to challenges in using
	them in practice for example they make obvious mistakes, output CPDAGs
	that need to be oriented using expert knowledge before estimation
	methods can be applied. As a result of these challenges, researchers
	typically construct these model by hand. In this paper, inspired by
	specification search in SEMs where the method suggests changes in the
	model that improves the fit the most, we present a method based on
	Conditional Independence testing and a novel measure of partial
	association. The method iteratively suggests addition and removal of
	edges between pairs of variables, while the user can choose the
	direction of the edge. Using empirical analysis we show that this
	manual construction approach can outperform automated algorithms if the
	user gets the direction correct in <x out of y> cases. We also show
	that the user expert can be replaced with an LLM.
\end{abstract}

>>>>

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Utilizing Expert Knowledge in Model Modification}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
	\emph{Causal Discovery} has received significant attention in the
	Directed Acyclic Graphs (DAGs) literature, leading to the development
	of numerous automated algorithms for constructing DAGs from data.
	Despite these advancements, the application of these algorithms in
	applied domains remain limited, with researchers often preferring to
	construct DAGs manually based on their expert knowledge. We hypothesize
	that the reluctance to use automated algorithms arises due to several
	practical issues, such as difficulty in choosing the best algorithm and
	their hyperparameters for a given dataset, algorithms can make obvious
	mistakes, and they output the Markov Equivalence Class (MEC) instead of
	a DAG, which is often required for further analysis. Even when using
	these algorithms, researchers must rely on their expert knowledge to
	make modifications, or orient edges of the learned model. Given this
	reliance on manual model construction and modification, in this paper,
	we present a method inspired by \emph{specification search} in
	Structural Equation Models (SEMs) to assist researchers in iteratively
	modifying their models. Specifically, we utilize implied Conditional
	Independence (CI) based model testing combined with a novel measure of
	partial association for mixed data to rank violations of implied CIs in
	data. Using this ranking users can make modifications to their models
	by either removing edges or adding edges between pair of variables
	while deciding the direction of the edge based on their domain
	expertise. The ranking allows users to prioritize the worst offending
	violations first. Through empirical analysis, when applied to learning
	the model, we demonstrate that taking a greedy approach (i.e., fixing
	the worst violation first) this approach can get comparable performance
	to automated algorithms if the users are able to get the edge direction
	correctly with an accuracy higher than $ 0.4 $. (check if this is
	correct). Alternatively, we show that a Large Language Model (LLM) can
	also be used to help in deciding the direction between the variables.
\end{abstract}
